18th International Satisfiability Modulo Theories Competition (SMT-COMP 2023): Rules and Procedures

### 2023-03-16 rules/rules23.tex
Updated affiliations in the new rule file.
SMT-COMP23 rules: what's new and proof track

\author{%
Fran\c{c}ois Bobot\\
CEA List \\
France \\
{\small\href{mailto:Francois.BOBOT@cea.fr}{\texttt{https://github.com/bobot}}} \\
\and
Martin Bromberger\\
TODO\\
Germany\\
{\small\href{mailto:TODO}{\texttt{TODO}}}\\
\and
Jochen Hoenicke\\
Certora\\
Germany\\
{\small\href{mailto:jochen@certora.com}{\texttt{jochen@certora.com}}}
}

@@

As in previous years, we have revised the rules slightly.  Some rule
changes in 2022 are designed to improve how logics are organized into
divisions. We have also added more divisions to the \mvaltrack and have defined a
new \prooftrack.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf Changes in divisions.} Last year we changed
  divisions to be composed of one or more related logics.
  %
  This year we will keep this organization but we will separate logics into
  divisions slightly differently: we will have a QF\_Datatypes division with the
  logics QF\_DT, QF\_UFDT and the QF\_Equality division now only contains the
  QF\_UF and QF\_AX logics.

  \rationale{The feedback from the participants in last year's competition was
    that datatypes were too dissimilar from QF\_UF and QF\_AX to be organized in
    the same division.}

  \item {\bf Logics can be non-competitive in competitive divisions.} Last year
  we considered every logic within a division to competitive unless the division
  as whole was non-competitive. This year if a logic is non-competitive within a
  division (i.e., only one competitive solver enrolled in it), it will not be
  ran.

  \rationale{We believe the previous setup could benefit too much a solver that
    supports more logics that its competitor.
    %
    In particular there was a whose winner was defined by the benchmarks of a
    non-competitive logic, which we believe is not ideal.
    %
    Note that the incentive for solvers to support more logics continues since
    as soon as there is more than one solver supporting a logic it will be
    competitive and the solver that does not support it will be disadvantaged.
  }

  \item {\bf \mvaltrack.}  \emergencystretch.5cm Last year's competition introduced the experimental
  divisions QF\_Equality (only with QF\_UF benchmarks), QF\_Equality+LinearArith (only with
  QF\_UFIDL, QF\_UFLIA, and QF\_UFLRA benchmarks), and QF\_Equality+Bitvec (only with
  QF\_UFBV benchmarks).
  %
  This year these divisions will no longer be experimental.
  %
  In addition, we add this year the new experimental division QF\_FPArith
  (excluding the logics QF\_ABVFP, QF\_AUFBVFP, and QF\_ABVFPLRA).

  \rationale{Last year's divisions ran successfully, so there
    is no reason to keep them experimental.
    %
    For the new divisions, as before, given the inconsistencies across different
    model producing solvers, we proceed in an experimental fashion to push for
    model standardization. We exclude logics containing arrays since their model
    production adds further complications which we only intend to tackle in the
    future.}

  \item {\bf \prooftrack.}  This track will be introduced this year.
  %
  Teams can submit proof-producing solvers together with proof checkers for
  their proof formats. The checker and format do not need to be from the same
  team as the solver.
  %
  We will compile and present the results, as well as assemble a panel of
  non-organizer experts to do a qualitative assessment for each proof-producing
  solver, proof format, and proof checker.
  %
  The solver, proof format and checker should be described in a system
  description (total of 4 pages rather than 2).

  \rationale{Given the many competing visions on how SMT proofs or a proof
    competition should look like, we decided to be less prescriptive and more
    descriptive for how SMT-COMP will approach proofs.
    %
    We will host a ``proof exhibition'' rather than a ``proof competition''
    track. The format would be with teams submitting both a proof-producing SMT
    solver and a proof checker for its proofs.
 }

@@

After the first line, the proof checker can provide arbitrary
\texttt{key=value} pairs.
%
These will be collected in the final result table and can include for example
information about the number of proof steps, the number of nodes in the proof,
the number of holes, the number of different proof rules, etc.

@@

@@

\header{Wrapper and Post-Processor.}
Solvers in this track will be wrapped in the competition to include the script
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/starexec_run_proof}}, which calls the solver and checker script.
The track will use
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/process}}
as a post-processor
to validate and accumulate the results.

@@

\item \emph{Cap the number of instances in a division.}
  The number of benchmarks in a division based on the size of the
  corresponding logics in SMT-LIB will be limited as follows.
  Let $n$ be the number of benchmarks in an SMT-LIB logic, then
  \begin{enumerate}
  \vspace{-1ex}
    \item \label{bench-sel-300} if $n \le 300$, all instances will be selected;
  \item \label{bench-sel-600} if $300 < n \leq 600$, a subset of 300 instances
    from the logic will be selected;
  \item \label{bench-sel-more} and if $n > 600$,
      50\% of the benchmarks of the logic will be selected.
  \end{enumerate}
\end{enumerate}
%
The selection process in cases \ref{bench-sel-600} and \ref{bench-sel-more}
above will guarantee the inclusion of new benchmarks by first picking randomly
one benchmark from each new benchmark family.  The rest of the benchmarks will
be chosen randomly from the remaining benchmarks using a uniform distribution.
%
The benchmark selection script will be publicly available at
\url{https://github.com/SMT-COMP/smt-comp/tree/master/tools} and
will use the same random seed as the rest of the competition.  The set of
benchmarks selected for the competition will be published when the competition
begins.

>>>>

\author{%
Fran\c{c}ois Bobot\\
CEA List \\
France \\
{\small\href{mailto:Francois.BOBOT@cea.fr}{\texttt{https://github.com/bobot}}} \\
\and
Martin Bromberger\\
Max Planck Institute for Informatics\\
Saarland Informatics Campus\\
Germany\\
{\small\href{mailto:mbromber@mpi-inf.mpg.de}{\texttt{mbromber@mpi-inf.mpg.de}}}\\
\and
Jochen Hoenicke\\
Certora\\
Germany\\
{\small\href{mailto:jochen@certora.com}{\texttt{jochen@certora.com}}}
}

@@

As in previous years, we have revised the rules slightly.  Some rule
changes in 2022 are designed to improve how logics are organized into
divisions. We have also added more divisions to the \mvaltrack and have defined a
new \prooftrack.
%
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf System description must be submitted with preliminary
      solver.} \todo{what}
%
  \rationale{\todo{reason}}

  \item {\bf \mvaltrack.}  \emergencystretch.5cm Last year's
    competition introduced the experimental divisions QF\_FPArith
    (excluding the logics QF\_ABVFP, QF\_AUFBVFP, and QF\_ABVFPLRA).
  %
  This year these divisions will no longer be experimental.
  %
  In addition, we add this year the new experimental division for
  non-linear arithmetic and for datatypes.  These include the divisions
  QF\_NonLinearIntArith, QF\_NonLinearRealArith,
  QF\_Equality+NonLinearArithm,
  and QF\_Datatypes
  (excluding all logics involving Arrays).  \todo{root-obj,
    partially defined functions}
%
  \rationale{Last year's divisions ran successfully, so there
    is no reason to keep them experimental.
    %
    For the new divisions, as before, given the inconsistencies across different
    model producing solvers, we proceed in an experimental fashion to push for
    model standardization. We exclude logics containing arrays since
    there is no standardized way of expressing array model values and
    we do not want to add too many new logics.}

\item {\bf \prooftrack.}
  Last year we added this as experimental track. We will continue with
  this track in more or less the same format.  We will reduce the
  number of benchmarks, though.
%  
  \rationale{We received several submissions that showed that there is
    interest in a proof track.

    We are still far away from a common proof format, so changing the
    track to a competition track or making it non-experimental is not
    feasible at this time.

    The number of benchmarks and the size of the produced proofs
    turned out to be a major stress to the StarExec infrastructure,
    while not giving only little insight.  Therefore, we will reduce
    the number of benchmarks in this track.
  }

@@

After the first line, the proof checker can provide arbitrary
\texttt{key=value} pairs.
%
These will be collected in the final result table and can include for example
information about the number of proof steps, the number of nodes in the proof,
the number of holes, the number of different proof rules, etc.
Each proof checker can define its own key/value pairs, but we suggest
to implement the following:
\begin{itemize}
\item \texttt{holes} -- number of unchecked axioms or rewrite rules (that
  can introduce unsoundness),
\item \texttt{axioms} -- number of checked axioms or rewrite rules,
\item \texttt{steps} -- number of proof steps (in some proof-specific
  granularity),
\item \texttt{assertions} -- number of asserted formulas used in the proof.
\end{itemize}

@@

\header{Evaluation Criteria.}
%
Since different proof formats are not easily comparable we will not do
any scoring by counting solved benchmarks.  Instead we will evaluate
both solvers and proof checkers using different criteria:
\begin{itemize}
\item \emph{soundness}: does the proof checker only accept correct
  proofs?  In particular, does it check that the proof proves the formula
  in the benchmark, and does it warn about unchecked rewrite rules?
\item \emph{trust-base}: how big is the trusted code and how much is
  shared between proof checker and solver?
\item \emph{proof overhead}: what is the performance of the solver and
  the proof checker compared to running the solver without proof
  production?
\item \emph{granularity}: what is the granularity of the proof?
\item \emph{readability}: is the proof easy to digest for a human?
\item \emph{extractability}: is the proof suitable for automatic
  processing?  Are there some implicit rewrites that one has to be
  aware of?
\item \emph{stability}: Is the proof checker specific to the current
  version of the solver and needs to be updated with each solver
  release, or is the proof checker fixed (except when adding new
  theories)?
\end{itemize}

@@

\header{Wrapper and Post-Processor.}
Solvers in this track will be wrapped in the competition to include the script
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/starexec_run_proof}},
which calls the solver and checker script.
Participants are encouraged to wrap their solvers themselves to test
their submission.
The track will use
{\url{https://github.com/SMT-COMP/postprocessors/tree/master/proof-track/process}}
as a post-processor
to validate and accumulate the results.

@@

\item \emph{Cap the number of instances in a division.}
  The number of benchmarks in a division based on the size of the
  corresponding logics in SMT-LIB will be limited as follows.
  Let $n$ be the number of benchmarks in an SMT-LIB logic, then
  the number of chosen benchmarks is $min(n, max(300, 50n/100))$:
  \begin{enumerate}
  \vspace{-1ex}
    \item \label{bench-sel-300} if $n \le 300$, all instances will be selected;
  \item \label{bench-sel-600} if $300 < n \leq 600$, a subset of 300 instances
    from the logic will be selected;
  \item \label{bench-sel-more} and if $n > 600$,
      50\% of the benchmarks of the logic will be selected.
    \end{enumerate}
    For the \prooftrack, we will only run $min(n, max(150,
    25n/100)))$ benchmarks.
\end{enumerate}
%
The selection process in cases \ref{bench-sel-600} and \ref{bench-sel-more}
above will guarantee the inclusion of new benchmarks by first picking randomly
one benchmark from each new benchmark family.  The rest of the benchmarks will
be chosen randomly from the remaining benchmarks using a uniform distribution.
%
The benchmark selection script will be publicly available at
\url{https://github.com/SMT-COMP/smt-comp/tree/master/tools} and
will use the same random seed as the rest of the competition.  The set of
benchmarks selected for the competition will be published when the competition
begins.
